---
title: "Stability Assignment"
author: "Mia Guarnieri, Lauren Harris"
date: "2023-06-08"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: paper
    code_folding: hide
    smooth_scroll: yes
    collapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages
```{r}
library(here)
library(tidyverse)
library(deSolve)
```


## Test the function

Definition of stability: a population that has a predator population size within 30% of the prey population size after 250 time steps.

## Varied hunting threshold

```{r}
#source function
source(here("stability", "lotvmodK.R"))

#use a wrapper function with pmap

#threshold list
thresh_l <- seq(from = 10, to = 150, by = 1)

#wrapper function
h_wrapper <- function(t, func, pop, rprey, alpha, eff, pmort, K, harvest, thresh) {
    pars <- list(rprey=rprey, alpha=alpha, eff=eff, pmort=pmort, K=K, harvest = harvest, thresh = thresh)
    
    #run the ode solver
    result <- ode(y = pop, times = times, func = func, parms = pars, maxsteps = 500000) 
  
    #turn into dataframe and get max value
    result <- as.data.frame(result)
    colnames(result) = c("time", "prey", "pred")
    final_pop <- tail(result, n = 1)
    
    return(final_pop)
}

#use pmap to apply wrapper to all parameter sets

times <- 1:250

pars_l <- list(rprey=0.95, alpha=0.01, eff=0.6, pmort=0.4, K=2000, harvest = 0.10, thresh = thresh_l)

pars_df <- as.data.frame(pars_l)

#setting initial population values
predi <- with(pars_l, pmort/(eff*alpha))
preyi <- with(pars_l, rprey/alpha*(1-predi/K))
pop <- c(prey = as.integer(preyi), pred = as.integer(predi))

all_results <- pars_df %>% 
  pmap(h_wrapper, t = times, func = lotvmodK, pop = pop)

all_results_df <- dplyr::bind_rows(all_results, .id = "pars") %>% 
  cbind(as.data.frame(thresh_l)) %>% 
  filter(time == 250.00) %>% 
  mutate(diff = ((abs(prey-pred))/prey),
         stable = ifelse(diff < 0.3, "true", "false"))
```

### Varied hunting

```{r}
#use a wrapper function with pmap

#harvest list
harvest_l <- seq(from = 0.05, to = 0.50, by = 0.01)

#use pmap to apply wrapper to all parameter sets

times <- 1:250

pars_l_h <- list(rprey=0.95, alpha=0.01, eff=0.6, pmort=0.4, K=2000, harvest = harvest_l, thresh = 61)

pars_df_h <- as.data.frame(pars_l_h)

#setting initial population values
predi <- with(pars_l, pmort/(eff*alpha))
preyi <- with(pars_l, rprey/alpha*(1-predi/K))
pop <- c(prey = as.integer(preyi), pred = as.integer(predi))

all_results_h <- pars_df_h %>% 
  pmap(h_wrapper, t = times, func = lotvmodK, pop = pop)

all_results_df_h <- dplyr::bind_rows(all_results_h, .id = "pars") %>% 
  cbind(as.data.frame(harvest_l)) %>% 
  filter(time == 250.00) %>% 
  mutate(diff = ((abs(prey-pred))/prey),
         stable = ifelse(diff < 0.3, "true", "false"))
```

## Plot final ode run

```{r}
#setting initial population values
predi <- with(pars_l, pmort/(eff*alpha))
preyi <- with(pars_l, rprey/alpha*(1-predi/K))
pop <- c(prey = as.integer(preyi), pred = as.integer(predi))

#selected parameters
pars_graph <- list(rprey=0.95, alpha=0.01, eff=0.6, pmort=0.4, K=2000, harvest = 0.13, thresh = 61)

#run the ode
graph_result <- ode(y = pop, times = times, func = lotvmodK, parms = pars_graph, maxsteps = 500000)

plot_df <- as.data.frame(graph_result) %>% 
  pivot_longer(-time, names_to = "animal", values_to = "pop")

ggplot(plot_df, aes(time, pop, col=animal)) + 
  geom_line() + 
  labs(x = "Time (days)", 
       y = "Population")
```

## Discussion

Our goal was to prevent the predator population from being too much larger than the prey population in the ecosystem. Therefore, our stability metric is whether the predator population size is within 30% of the prey population size after 250 time steps, with the goal of keeping predator and prey sizes as close together as possible to ensure that both populations can remain stable. When we ran our ode solver, we determined that the ideal hunting threshold is 61 prey individuals; below this level, hunting should not occur. We also determined that the ideal hunting rate is 0.13, or 13% of the prey population. These were the variables that kept the predator and prey populations as close together as possible given the starting parameters.
